Nitric oxide feedback to ciliary photoreceptor cells gates a UV avoidance circuit

### 2024-02-05 Manuscript/Manuscript.Rmd
MS, version info

---
title: Nitric oxide feedback to ciliary photoreceptor cells gates a UV avoidance circuit
author: 
  - name: "Kei Jokura<sup>1,`*`</sup>, Nobuo Ueda<sup>2</sup>, Martin Gühmann<sup>3</sup>, Luis Alfonso Yañez-Guerra<sup>1,4</sup>, Piotr Słowiński<sup>1</sup>, Kyle C. A. Wedgwood<sup>1</sup> and Gáspár Jékely<sup>1,5,`*`</sup>"
    affiliation: "<sup>1</sup>Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4 4QD, United Kingdom <sup>2</sup>Okinawa Institute of Science and Technology, Okinawa, Japan, <sup>3</sup>School of Biological Sciences, University of Bristol, Bristol, United Kingdom, <sup>4</sup>School of Biological sciences, University of Southampton, Southampton, United Kingdom, <sup>5</sup>Centre for Organismal Studies (COS), University of Heidelberg, 69120 Heidelberg, Germany  <br> <sup>*</sup>Correspondence`:` k.jokura @ exeter.ac.uk, gaspar.jekely @ cos.uni-heidelberg.de"
output:
  pdf_document:
    latex_engine: lualatex
    template: default
  html_document:
    theme: lumen
  word_document: default
geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.5in
mainfont: Arial
monofont: Times New Roman
documentclass: article
classoption: onecolumn
linestretch: 1.2
fontsize: 10pt
bibliography: references.bib
csl: elife.csl
editor_options: 
  markdown: 
    wrap: sentence
---

@@

```{r echo=FALSE,fig.cap="**Figure 3. NOS is required for UV avoidance in *Platynereis* larvae.**  (**A**) Swimming trajectories of wild type (WT, n=32) and *NOS* mutant (*NOSΔ11/Δ11*, n=26 and *NOSΔ23/Δ23*, n=47) three-day-old larvae. All trajectories start at 0 x and y position and time 0 corresponding to 10 sec before the onset of 395 nm stimulation from the side. (**B**) Vertical position of batches of wild type and mutant three-day-old larvae over time under 395 nm UV stimulation. The starting position of each larval trajectory was set to 0. (**C**) Comparison of vertical position of wild type and mutant three-day-old larvae after 30 seconds of UV stimulation. One-way ANOVA with Dunnett’s multiple-comparison test are shown. The wild type (WT, n=32) and *NOS* mutant (*NOSΔ11/Δ11*, n=26 and *NOSΔ23/Δ23*, n=47) larvae. (**D**) Vertical displacement in 30 sec bins of wild type (10 batches) and mutant (*NOSΔ11/Δ11*, 5 batches and *NOSΔ23/Δ23*, 7 batches) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top. One-way ANOVA with Dunnett’s multiple-comparison test are shown. *P < 0.05. (**E**) Vertical position of batches of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae over time under 395 nm UV stimulation. The starting position of each larval trajectory was set to 0. (**F**) Comparison of vertical position of control (n=6) and L-NAME-treated (0.1 and 1 mM, n=6 each) three-day-old larvae after 30 seconds of UV stimulation. One-way ANOVA with Dunnett’s multiple-comparison test are shown. (**G**) Vertical displacement in 30 sec bins of control (15 batches) and L-NAME-treated (0.1 and 1 mM, 15 batches each) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top. One-way ANOVA with Dunnett’s multiple-comparison test are shown. *P < 0.05. <br> Figure 3 -- source data 1-7. Source data for panels A-G."}

@@

```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 3. Expression of NIT-GC1 and NIT-GC2 in the cPRCs** (**A**) HCR co-expression analysis of *NIT-GC1* (magenta) and *MLD/pedal-peptide-2 proneuropeptide* (green). Anterior view of a two-day-old larva. Nuclei were stained with DAPI (cyan). **(B, C)** Expression of *NIT-GC2* (magenta) detected by in situ HCR. Anterior (B) and ventral (C) views of a three-day-old larva. Nuclei were stained with DAPI (cyan). **(D, E)** Immunostaining for NIT-GC1 (D) and NIT-GC2 (E) antibodies in larvae injected with NIT-GC1 (D) or NIT-GC2 (E) morpholinos. Acetylated α-tubulin staining (green) highlights the neuropil and cPRC cilia. Anterior view of a two-day-old larva. (**F**) The fluorescence intensities of the NIT-GC1 or NIT-GC2 in the control and morphant (MO1 and MO2) larvae (n=6 each) were normalized to the Acetylated α-tubulin signals and plotted as a relative fluorescence. One-way ANOVA with Dunnett’s multiple-comparison test; p-values (‘***’ <0.001) are shown. <br> Figure 4 -- figure suppmenet 3 -- source data 1. Comparison NIT-GCs signals readings."}
knitr::include_graphics(path = "figures/Fig4_sup3.png")

>>>>

---
title: Nitric oxide feedback to ciliary photoreceptor cells gates a UV avoidance circuit
author: 
  - name: "Kei Jokura<sup>1,`*`</sup>, Nobuo Ueda<sup>2</sup>, Martin Gühmann<sup>3</sup>, Luis Alfonso Yañez-Guerra<sup>1,4</sup>, Piotr Słowiński<sup>1</sup>, Kyle C. A. Wedgwood<sup>1</sup> and Gáspár Jékely<sup>1,5,`*`</sup>"
    affiliation: "<sup>1</sup>Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4 4QD, United Kingdom <sup>2</sup>Okinawa Institute of Science and Technology, Okinawa, Japan, <sup>3</sup>School of Biological Sciences, University of Bristol, Bristol, United Kingdom, <sup>4</sup>School of Biological sciences, University of Southampton, Southampton, United Kingdom, <sup>5</sup>Centre for Organismal Studies (COS), University of Heidelberg, 69120 Heidelberg, Germany  <br> <sup>*</sup>Correspondence`:` k.jokura @ exeter.ac.uk, gaspar.jekely @ cos.uni-heidelberg.de"
output:
  html_document:
    theme: lumen
  pdf_document:
    latex_engine: lualatex
    template: default
  word_document: default
geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.5in
mainfont: Arial
monofont: Times New Roman
documentclass: article
classoption: onecolumn
linestretch: 1.2
fontsize: 10pt
bibliography: references.bib
csl: elife.csl
editor_options: 
  markdown: 
    wrap: sentence
---

@@

```{r echo=FALSE,fig.cap="**Figure 3. NOS is required for UV avoidance in *Platynereis* larvae.**  (**A**) Swimming trajectories of wild type (WT, n=32) and *NOS* mutant (*NOSΔ11/Δ11*, n=26 and *NOSΔ23/Δ23*, n=47) three-day-old larvae. All trajectories start at 0 x and y position and time 0 corresponding to 10 sec before the onset of 395 nm stimulation from the side. (**B**) Vertical position of batches of wild type and mutant three-day-old larvae over time under 395 nm UV stimulation. The starting position of each larval trajectory was set to 0. (**C**) Comparison of vertical position of wild type and mutant three-day-old larvae after 30 seconds of UV stimulation. One-way ANOVA with Dunnett’s multiple-comparison test are shown. The wild type (WT, n=32) and *NOS* mutant (*NOSΔ11/Δ11*, n=26 and *NOSΔ23/Δ23*, n=47) larvae. (**D**) Vertical displacement in 30 sec bins of wild type (10 batches) and mutant (*NOSΔ11/Δ11*, 5 batches and *NOSΔ23/Δ23*, 7 batches) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top. One-way ANOVA with Dunnett’s multiple-comparison test are shown. *P < 0.05. (**E**) Vertical position of batches of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae over time under 395 nm UV stimulation. The starting position of each larval trajectory was set to 0. (**F**) Comparison of vertical position of control (n=6) and L-NAME-treated (0.1 and 1 mM, n=6 each) three-day-old larvae after 30 seconds of UV stimulation. One-way ANOVA with Dunnett’s multiple-comparison test are shown. (**G**) Vertical displacement in 30 sec bins of control (15 batches) and L-NAME-treated (0.1 and 1 mM, 15 batches each) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top. One-way ANOVA with Dunnett’s multiple-comparison test are shown. *P < 0.05, **P < 0.01. <br> Figure 3 -- source data 1-7. Source data for panels A-G."}

@@

```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 3. Expression of NIT-GC1 and NIT-GC2 in the cPRCs** (**A**) HCR co-expression analysis of *NIT-GC1* (magenta) and *MLD/pedal-peptide-2 proneuropeptide* (green). Anterior view of a two-day-old larva. Nuclei were stained with DAPI (cyan). **(B, C)** Expression of *NIT-GC2* (magenta) detected by in situ HCR. Anterior (B) and ventral (C) views of a three-day-old larva. Nuclei were stained with DAPI (cyan). **(D, E)** Immunostaining for NIT-GC1 (D) and NIT-GC2 (E) antibodies in larvae injected with NIT-GC1 (D) or NIT-GC2 (E) morpholinos. Acetylated α-tubulin staining (green) highlights the neuropil and cPRC cilia. Anterior view of a two-day-old larva. (**F**) The fluorescence intensities of the NIT-GC1 or NIT-GC2 in the control and morphant (MO1 and MO2) larvae (n=6 each) were normalized to the acetylated α-tubulin signals and plotted as a relative fluorescence. One-way ANOVA with Dunnett’s multiple-comparison test; p-values (‘***’ <0.001) are shown. <br> Figure 4 -- figure suppmenet 3 -- source data 1. Comparison NIT-GCs signals readings."}
knitr::include_graphics(path = "figures/Fig4_sup3.png")
